---
title: "Ants_Infomap"
output:
  pdf_document: default
  html_document: default
---

```{r setup}
# Install (if not installed) and load necessary packages
package.list=c("attempt", "cowplot", "igraph", "ggalluvial", "magrittr", "metafolio", "tidyverse", "vegan", "devtools")

loaded <- package.list %in% .packages()
package.list <- package.list[!loaded]
installed <- package.list %in% .packages(TRUE)
if(!all(installed)) install.packages(package.list[!installed], repose="http://cran.rstudio.com/")

#Install Infomapecology
devtools::install_github('Ecological-Complexity-Lab/infomap_ecology_package', force=T)
library(infomapecology)

# Check the version. Should be at least 0.1.1.1
packageDescription('infomapecology')

# Install infomap if you have not done so externally (see previous section in this readme)
setwd('/home/sshatzkin/Networks in Biology')
install_infomap()

# Check Infomap is running
setwd('/home/sshatzkin/Networks in Biology')
check_infomap() #Make sure file can be run correctly. Should return True
```

```{r}
library(igraph)
library(bipartite)
library(tidyverse)
library(magrittr)
library(infomapecology)
library(readxl)
library(ggalluvial)
```

# 1. Infomap Analysis
### Importing the Data
```{r}
numFiles <- 41

import_data <- function (string_format, numFiles){
  # Create the file names
  file_names <- sprintf(string_format, 1:numFiles);
  
  # Create an empty linked list
  colony_data <- vector(mode = "list", length = numFiles)
  
  # Read in all of the files
  for (i in 1:numFiles){
    colony_data[[i]] <- read_graph(file_names[i], format = "graphml")
  }
  
  return (colony_data)
}

colony1_data <- import_data('./ants_proximity_weighted/ant_mersch_col1_day%02d_attribute.graphml', numFiles)


for (i in 1:2){
  plot.igraph(colony1_data[[i]]) # This plot is not particularly informative
}

```

### Perparing the data.
The data must be converted to an infomap compatible format

#### Helper functions
```{r}
# Function that takes a vertex list and the number of layers in the multilayer network, and produces a formatted df containing the multilayer edges in that network (with edges of weight 1)
interlayer_edge_list <- function (vertex_names, num_layers){
  
  numVerts <- length (vertex_names)
  numRows <- numVerts * (num_layers - 1)
  
  col_layer_from <- vector(length = numRows)
  col_layer_to <- vector(length = numRows)
  col_node_from <- vector(length = numRows)
  col_node_to <- vector(length = numRows)
  col_weights <- vector(length = numRows)
  
  for (l in 1:(num_layers-1)){
    for (vertex in 1:numVerts){
      currIndex <- ((l-1) * numVerts) + vertex 
      col_layer_from[currIndex] <- l
      col_layer_to[currIndex] <- l + 1
      col_node_from[currIndex] <- vertex_names[vertex]
      col_node_to[currIndex] <- vertex_names[vertex]
      col_weights[currIndex] <- 1
    }
  }
  
  df <- data.frame("layer_from" = col_layer_from, "node_from" = col_node_from, "layer_to" = col_layer_to, "node_to" = col_node_to, "weight" = col_weights)
  
  return (df)
}

# Joins a list of dataframes into a single dataframe 
join_edge_dfs <- function(edge_dfs, num_lists){
  df <- edge_dfs[[1]]
  for (i in 2:num_lists){
    df <- rbind(df, edge_dfs[[i]])
  }
  return(df)
}


```

#### Main datatype conversion function
```{r}
# Now we need to convert this igraph data into data structures that infomap can take

igraphs_to_multilayer_df <- function (igraph_list, num_elements){
  
  # Create empty lists
  colony_edges <- vector(mode = "list", length = num_elements)
  colony_nodes <- vector(mode = "list", length = num_elements)

  # Loop over layers
  for (i in 1:num_elements){
    
    # Pull out the edge lists from each layer
    colony_edges[[i]] <- data.frame(cbind(get.edgelist(igraph_list[[i]]), round (E(igraph_list[[i]])$weight, 3)))
    
    #Assign column labels
    names(colony_edges[[i]]) <- c('node_from', 'node_to', 'weight')
    

    #Generate an array of layer labels
    layer_arr <- rep(i, nrow(colony_edges[[i]]))
    
    colony_edges[[i]]$layer_from <- layer_arr
    colony_edges[[i]]$layer_to <- layer_arr
    
    # Reorder the columns
    colony_edges[[i]] <- colony_edges[[i]][,c('layer_from', 'node_from', 'layer_to', 'node_to', 'weight')]
    
    
    
    # Pull out the vertices from each layer
    dayi_vertices <- V(igraph_list[[i]])
    
    # Prepare the node lists
    colony_nodes[[i]] <- data.frame(as_ids(dayi_vertices))
    names(colony_nodes[[i]]) <- c('node_id')
    
  }
  
  
  # Construct inter-layer edges
  interlayer_edges <- interlayer_edge_list(colony_nodes[[i]]$node_id, num_elements);
  

  # Now merge these smaller lists to form the full edge and node lists
  colony_edges_joined <- join_edge_dfs(colony_edges, num_elements)
  colony_edges_joined <- rbind(colony_edges_joined, interlayer_edges)
  
  
  return (list(colony_edges_joined, colony_nodes[[1]]))
  
}


```


### Plot all 41 days as one multilayer network
```{r}

# Function takes the igraph_list, converts to an infomap data structure, and returns the module results
igraph_to_infomap_modules <- function (igraph_list, num_layers){
  
  # Convert the igraph list into a dataframe with interlayer edges
  colony_dfs <- igraphs_to_multilayer_df(igraph_list, num_layers)

  colony_edges <- colony_dfs[[1]]
  colony_nodes <- colony_dfs[[2]]

  # Build layer_map
  layer_map <- tibble(layer_id=1:num_layers, layer_name=1:num_layers)

  # Create the multilary object
  colony <- create_multilayer_object(extended = colony_edges, nodes = colony_nodes, intra_output_extended = T, inter_output_extended = T, layers = layer_map)

  # Run infomap
  colony_modules <- run_infomap_multilayer(M=colony, relax = F, flow_model = 'directed', silent = T, trials = 100, seed = 497294, temporal_network = T)
  
  return (colony_modules)  
}

```


```{r}
colony1_modules <- igraph_to_infomap_modules(colony1_data, numFiles)
colony1_modules
```

```{r}
plot_persistance <- function (module_data, num_layers, num_modules){
  # Module Persistance
  module_persistence <- module_data$modules %>%
  group_by(module) %>%
  summarise(b=min(layer_id), d=max(layer_id), persistence=d-b+1)

  # Plot modules' persistence
  plot_multilayer_modules(module_data, type = 'rectangle', color_modules = T)+
    scale_x_continuous(breaks = seq(0, num_layers, 5))+
    scale_y_continuous(breaks = seq(0, num_modules, 2))+
    scale_fill_viridis_c()+
    theme_bw()+
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.title = element_text(size = 20),
          legend.text = element_text(size=15),
          legend.title = element_text(size=20))
}

plot_persistance(colony1_modules, 41, 20)


```
```{r}
plot_alluvial <- function (module_data, num_layers, num_ants){
  # Plot species flow through modules in time
  plot_multilayer_alluvial(module_data, module_labels = F)+
  scale_x_continuous(breaks=seq(0, num_layers, 5))+
  scale_y_continuous(breaks=seq(0,num_ants,20))+
  labs(y='Number of Ants')+
  theme_bw()+
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text = element_text(color='black', size = 20),
        axis.title = element_text(size=20))
}

plot_alluvial(colony1_modules, 41, 120)

```
#### Test with another colony
```{r}

# Import all of the colony2 data files
colony2_data <- import_data('./ants_proximity_weighted/ant_mersch_col2_day%02d_attribute.graphml', numFiles)

# Convert a list of igraphs into a processed infomap module
colony2_modules <- igraph_to_infomap_modules(colony2_data, numFiles)

# Plots functions based on just module data
plot_persistance(colony2_modules, 41, 20)
plot_alluvial(colony2_modules, 41, 120)


```

### Plot in 10-day increments
```{r}

divi_modules <- igraph_to_infomap_modules(colony1_data[1:11], 11)
plot_persistance(divi_modules, 11, 20)
plot_alluvial(divi_modules, 11, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[12:21], 10)
plot_persistance(divi_modules, 10, 20)
plot_alluvial(divi_modules, 10, 120)

```
```{r}
divi_modules <- igraph_to_infomap_modules(colony1_data[22:31], 10)
plot_persistance(divi_modules, 10, 20)
plot_alluvial(divi_modules, 10, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[32:41], 10)
plot_persistance(divi_modules, 10, 20)
plot_alluvial(divi_modules, 10, 120)


```
```{r}
divi_modules <- igraph_to_infomap_modules(colony1_data[1:6], 6)
plot_persistance(divi_modules, 6, 20)
plot_alluvial(divi_modules, 6, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[7:11], 5)
plot_persistance(divi_modules, 5, 20)
plot_alluvial(divi_modules, 5, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[12:16], 5)
plot_persistance(divi_modules, 5, 20)
plot_alluvial(divi_modules, 5, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[17:21], 5)
plot_persistance(divi_modules, 5, 20)
plot_alluvial(divi_modules, 5, 120)


divi_modules <- igraph_to_infomap_modules(colony1_data[32:36], 5)
plot_persistance(divi_modules, 5, 20)
plot_alluvial(divi_modules, 5, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[37:41], 5)
plot_persistance(divi_modules, 5, 20)
plot_alluvial(divi_modules, 5, 120)
```

#### Double Layer Networks
```{r}
divi_modules <- igraph_to_infomap_modules(colony1_data[1:2], 2)
plot_persistance(divi_modules, 2, 20)
plot_alluvial(divi_modules, 2, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[2:3], 2)
plot_persistance(divi_modules, 2, 20)
plot_alluvial(divi_modules, 2, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[3:4], 2)
plot_persistance(divi_modules, 2, 20)
plot_alluvial(divi_modules, 2, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[4:5], 2)
plot_persistance(divi_modules, 2, 20)
plot_alluvial(divi_modules, 2, 120)

divi_modules <- igraph_to_infomap_modules(colony1_data[5:6], 2)
plot_persistance(divi_modules, 2, 20)
plot_alluvial(divi_modules, 2, 120)
```

### Monolayer Approach
Note that this code is a template of the monlayer adaptation of this method, but is not complete.
```{r}
igraphs_to_monolayer_dfs <- function (igraph_list, num_elements){
  
  # Create empty lists
  #colony_layers <- vector(mode = "list", length = num_elements)

  # Loop over layers
  #for (i in 1:num_elements){
  #  adj_mtrx <- as_adjacency_matrix(igraph_list[[i]])
   
    # Pull out the edge lists from each layer
    #colony_layers[[i]] <- create_monolayer_object(x = adj_mtrx, directed = F, bipartite = F)
    
    #Assign column labels
    #names(colony_edges[[i]]) <- c('node_from', 'node_to', 'weight')
    
    
  #}
  
  #return (colony_layers)
}

#colony1_monolayers <- igraphs_to_monolayer_dfs(colony1_data, 41)
```